We Are All Trying to Describe Something And We Don't Have the Words

### 2023-04-12 outputs/paper/paper.qmd
new artists, new df

---
title: "We Are All Trying to Describe Something And We Don't Have the Words"
subtitle: "an exploration of the Jack Antonoffification of Pop Music"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: LINK."
date: today
abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API. Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an Antonoff-y song, and assess whether or not the audio features defined by Spotify can capture this. I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me. There will be a Bayesian inference model. It will be a logistic regression. We do not know whether or not it will be good. Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Midnight was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".

@@

library(lubridate)
df <- read.csv(here::here("inputs/data/df.csv"))
df$jack <- factor(df$jack)
df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

>>>>

---
title: "We Are All Trying to Describe Something And We Don't Have the Words"
subtitle: "An exploration of the Jack Antonoffification of Pop Music"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: LINK."
date: today
abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API. Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an Antonoff-y song, and assess whether or not the audio features defined by Spotify can capture this. I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me. There will be a Bayesian inference model. It will be a logistic regression. We do not know whether or not it will be good. Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Midnights was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".

@@

library(lubridate)
library(tidyverse)
df <- read_csv(here::here("inputs/data/df.csv"))
df$jack <- factor(df$jack)
df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080", "#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap( ~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
    geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
    geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
    geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080","#FFB6C1")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

I'm gonna fix the axis labels later

#| label: fig-quadvalence1
#| fig-cap: Valence and Energy Quadrant 
#| echo: false
#| warning: false

ggplot(data = df, aes(x = energy, y = valence)) + 
  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$valence), color = "gray") +
  annotate("text", x = mean(df$energy), y = max(df$valence), 
           label = "High Energy, High Valence") +
  annotate("text", x = mean(df$energy), y = min(df$valence), 
           label = "High Energy, Low Valence") +
  annotate("text", x = min(df$energy), y = mean(df$valence), 
           label = "Low Energy, High Valence") +
  annotate("text", x = max(df$energy), y = mean(df$valence), 
           label = "Low Energy, Low Valence")

#| label: fig-quadvalence2
#| fig-cap: Valence and Energy Quadrant
#| echo: false
#| warning: false

ggplot(data = df, aes(x = energy, y = valence)) + 
  geom_point(aes(color = artist_name), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$valence), color = "gray") +
  annotate("text", x = mean(df$energy), y = max(df$valence), 
           label = "High Energy, High Valence") +
  annotate("text", x = mean(df$energy), y = min(df$valence), 
           label = "High Energy, Low Valence") +
  annotate("text", x = min(df$energy), y = mean(df$valence), 
           label = "Low Energy, High Valence") +
  annotate("text", x = max(df$energy), y = mean(df$valence), 
           label = "Low Energy, Low Valence") 
```

#| label: fig-quaddanceability
#| fig-cap: Danceability and Energy Quadrant
#| echo: false
#| warning: false

ggplot(data = df, aes(x = energy, y = danceability)) + 
  geom_point(aes(color = artist_name), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$danceability), color = "gray") +
  annotate("text", x = mean(df$energy), y = max(df$danceability), 
           label = "High Energy, High Danceability") +
  annotate("text", x = mean(df$energy), y = min(df$danceability), 
           label = "High Energy, Low Danceability") +
  annotate("text", x = min(df$energy), y = mean(df$danceability), 
           label = "Low Energy, High Danceability") +
  annotate("text", x = max(df$energy), y = mean(df$danceability), 
           label = "Low Energy, Low Danceability")

#| label: fig-quaddanceability2
#| fig-cap: Danceability and Energy Quadrant
#| echo: false
#| warning: false

ggplot(data = df, aes(x = energy, y = danceability)) + 
  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$danceability), color = "gray") +
  annotate("text", x = mean(df$energy), y = max(df$danceability), 
           label = "High Energy, High Danceability") +
  annotate("text", x = mean(df$energy), y = min(df$danceability), 
           label = "High Energy, Low Danceability") +
  annotate("text", x = min(df$energy), y = mean(df$danceability), 
           label = "Low Energy, High Danceability") +
  annotate("text", x = max(df$energy), y = mean(df$danceability), 
           label = "Low Energy, Low Danceability")
